{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-12-14 sftrends_manu.Rnw
added trncomp figures

@@

@@

>>>>

Effects of flow on trend interpration shown in \cref{fig:trndcomp1,fig:trndcomp2,fig:trndcomp3}.

@@

<<trndcomp, echo = F, warning = F, message = F, cache = T>>=
data(trnds_seasyr)

# response labels 
reslabs <- list(
  shrt = c('din', 'nh', 'no23'),
  expr = c(
    'DIN', 
    expression(paste(NH [4] ^ '+')),
    expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-'))
  )
)

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, ztest, tau, perchg, ann, trndvar) %>% 
  filter(!trndvar %in% 'res') %>% 
  mutate(
    Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'P8', 'C10', 'C3')),
    ztest = ifelse(ztest < 0.05, '(< 0.05)', '(ns)'),
    resvar = factor(resvar, levels = reslabs$shrt, labels = reslabs$expr), 
    trndvar = factor(trndvar, levels = c('fits', 'norm'), labels = c('Pred.', 'Flow-norm.'))
  ) %>% 
  unite('trnd', trndvar, ztest, sep = ' ')

# ggplot theme
mythm <- theme_minimal() +
  theme(
    legend.position = 'top', 
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )

# year only
toplo1 <- filter(trnds_seasyr, cat %in% c('1976-1995', '1996-2013'))
p1 <- ggplot(toplo1, aes(x = tau, y = Site_Code)) + 
  geom_point(aes(pch = trnd, colour = abs(tau)), size = 4, alpha = 0.8) + 
  facet_grid(cat~resvar, labeller = label_parsed
    ) + 
  geom_vline(xintercept = 0, linetype = 'dashed') + 
  mythm + 
  xlab(expression(tau)) +
  scale_shape_manual(values = c(16, 1, 17, 2)) +
  scale_colour_gradient(low = 'grey60', high = 'black', guide = F) + 
  scale_x_continuous(limits = c(-1, 1)) + 
  geom_hline(yintercept = c(3.5, 6.5), colour = 'grey', alpha = 0.6)

# seasonal, first years
toplo2 <- filter(trnds_seasyr, !cat %in% c('1976-1995', '1996-2013')& ann == 'bef') 
p2 <- ggplot(toplo2, aes(x = tau, y = Site_Code)) + 
  geom_point(aes(pch = trnd, colour = abs(tau)), size = 4, alpha = 0.8) + 
  facet_grid(cat~resvar, labeller = label_parsed
    ) + 
  geom_vline(xintercept = 0, linetype = 'dashed') + 
  mythm + 
  xlab(expression(tau)) +
  scale_shape_manual(values = c(16, 1, 17, 2)) +
  scale_colour_gradient(low = 'grey60', high = 'black', guide = F) + 
  scale_x_continuous(limits = c(-1, 1)) + 
  geom_hline(yintercept = c(3.5, 6.5), colour = 'grey', alpha = 0.6)

# seasonal, first years
toplo3 <- filter(trnds_seasyr, !cat %in% c('1976-1995', '1996-2013')& ann == 'aft')
p3 <- ggplot(toplo3, aes(x = tau, y = Site_Code)) + 
  geom_point(aes(pch = trnd, colour = abs(tau)), size = 4, alpha = 0.8) + 
  facet_grid(cat~resvar, labeller = label_parsed
    ) + 
  geom_vline(xintercept = 0, linetype = 'dashed') + 
  mythm + 
  xlab(expression(tau)) +
  scale_shape_manual(values = c(16, 1, 17, 2)) +
  scale_colour_gradient(low = 'grey60', high = 'black', guide = F) + 
  scale_x_continuous(limits = c(-1, 1)) + 
  geom_hline(yintercept = c(3.5, 6.5), colour = 'grey', alpha = 0.6)

pdf('figs/trndcomp1.pdf', height = 5, width = 9, family = 'serif')
p1
dev.off()

pdf('figs/trndcomp2.pdf', height = 8, width = 9, family = 'serif')
p2
dev.off()

pdf('figs/trndcomp3.pdf', height = 8, width = 9, family = 'serif')
p3
dev.off()
@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,page=1]{figs/trndcomp1.pdf}
\caption{Results from seasonal Kendall tests on model predictions from \ac{wrtds}. Trends were evaluated using predicted (triangles) and flow-normalized (circles) results to identify potential effects of flow. Results are the estimated direction and magnitude of the trend ($\tau$, varying from -1 to 1 with significance at $\alpha = 0.05$). Trends are shown separately for different annual groupings. See \cref{fig:trndcomp2,fig:trndcomp3} for seasonal groupings.}
\label{fig:trndcomp1}   
\end{figure}

@@

\clearpage

% supplementary material
\beginsupplement

\begin{figure}
\centering
\includegraphics[width=1\textwidth,page=1]{figs/trndcomp2.pdf}
\caption{Results from seasonal Kendall tests on model predictions from \ac{wrtds}. Trends were evaluated using predicted (triangles) and flow-normalized (circles) results to identify potential effects of flow. Results are the estimated direction and magnitude of the trend ($\tau$, varying from -1 to 1 with significance at $\alpha = 0.05$). Trends are shown separately for different seasonal groupings from 1976-1995. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. See \cref{fig:trndcomp1} for annual comparisons.}
\label{fig:trndcomp2}   
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,page=1]{figs/trndcomp3.pdf}
\caption{Results from seasonal Kendall tests on model predictions from \ac{wrtds}. Trends were evaluated using predicted (triangles) and flow-normalized (circles) results to identify potential effects of flow. Results are the estimated direction and magnitude of the trend ($\tau$, varying from -1 to 1 with significance at $\alpha = 0.05$). Trends are shown separately for different seasonal groupings from 1996-2013. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. See \cref{fig:trndcomp1} for annual comparisons.}
\label{fig:trndcomp3}   
\end{figure}

### 2016-12-14 sftrends_manu.tex
added trncomp figures

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Dec 13 16:33:55 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Wed Dec 14 15:30:51 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

pdf 
  2 
pdf 
  2 
pdf 
  2

\begin{figure}
\centering
\includegraphics[width=1\textwidth,page=1]{figs/trndcomp1.pdf}
\caption{Results from seasonal Kendall tests on model predictions from \ac{wrtds}. Trends were evaluated using predicted (triangles) and flow-normalized (circles) results to identify potential effects of flow. Results are the estimated direction and magnitude of the trend ($\tau$, varying from -1 to 1 with significance at $\alpha = 0.05$). Trends are shown separately for different annual groupings. See \cref{fig:trndcomp2,fig:trndcomp3} for seasonal groupings.}
\label{fig:trndcomp1}   
\end{figure}
